The Evolution of Phenotypically Invariant Genetic Networks

### 2017-12-08 kalman_walk_paper.tex
minor fixing
more minor fixing
.

Let two points on $\allS$ be $x_1$ and $x_2$, let $\bar x = (x_1+x_2)/2$, and let $z=(x_2 - x_1)/2$.
Then with $\grad D$ and $\grad^2 D$ the first and second derivatives of $D$, respectively,
Taylor expanding about $x_1$ and $x_2$ finds that
\begin{align*}
    D(\bar x) 
    &= D(x_1) + \grad D(x_1) \cdot z + \frac{1}{2} z^T \grad^2  D(x_1) z + O(\|z\|^3) \\
    &=  D(x_2) - \grad D(x_2) \cdot z + \frac{1}{2} z^T \grad^2  D(x_2) z + O(\|z\|^3) .
\end{align*}
Now, since $ D(x_1) =  D(x_2) = \grad D(x_1) = \grad D(x_2) = 0$ and
\begin{align*}
    \grad D(x_2) = \grad D(x_1) + 2 z^T \grad^2  D(x_1) + O(\|z\|^2), \quad \text{and} \\
    \grad^2 D(x_2) = \grad^2 D(x_1) + O(\|z\|), 
\end{align*}
adding together the two equations above and dividing by two gets that
\begin{align*}
     D(\bar x) 
    &= \optph - \frac{3}{2} z^T \grad^2  D(x_1) z + O(\|z\|^3) .
\end{align*}

@@

We could use this to get the quadratic approximation to $D$ near the optimal set.
To do so, it'd be nice to have a way to compute the inner integral above.
Suppose that we diagonalize $A = U \Lambda U^{-1}$.
Then
\begin{equation} \label{eqn:exp_deriv}
  \begin{aligned}
      \int_0^t e^{As} Z e^{A(t-s)} ds 
      &=
      \int_0^t U e^{\Lambda s} U^{-1} Z U e^{\Lambda (t-s)} U^{-1} ds 
  \end{aligned}
\end{equation}
Now, notice that
\begin{equation}
  \begin{aligned}
      \int_0^t e^{s \lambda_i} e^{(t-s) \lambda_j} ds
      &=
      \frac{ e^{t \lambda_i} - e^{t \lambda_j} }{ \lambda_i - \lambda_j } .
  \end{aligned}
\end{equation}
Therefore, 
defining
\begin{equation}
    X_{ij}(t,Z) = \left( U^{-1} Z U \right)_{ij}
      \frac{ e^{t \lambda_i} - e^{t \lambda_j} }{ \lambda_i - \lambda_j } 
\end{equation}
moving the $U$ and $U^{-1}$ outside the integral and integrating we get that
\begin{equation}
  \begin{aligned}
      \int_0^t e^{As} Z e^{A(t-s)} ds 
      &=
      U X(t,Z) U^{-1} .
  \end{aligned}
\end{equation}

Following on from above, we see that if $Z=\Delta_{k \ell}$, then
\begin{equation}
  \begin{aligned}
      X_{ij}^{k\ell}(t) = 
      \frac{ e^{t \lambda_i} - e^{t \lambda_j} }{ \lambda_i - \lambda_j } 
      (U^{-1})_{\cdot k} U_{\ell \cdot},
  \end{aligned}
\end{equation}
where $U_{k \cdot}$ is the $k$th row of $U$,
and so
\begin{equation}
    \begin{aligned}
        H_{ij, k\ell}(A)
        &=
        \int_0^\infty
            \rho(t) \tr\left\{ C U X^{ij}(t) U^{-1} B B^T (U^{-1})^T X^{k\ell}(t)^T U^T C^T \right\}
        dt .
    \end{aligned}
\end{equation}
This implies that
\begin{equation}
    \begin{aligned}
        D(A_0+\epsilon Z)^2
        &\approx \epsilon^2\sum_{ijk\ell} H^{ij,k\ell} Z_{ij} Z_{k\ell} 
    \end{aligned}
\end{equation}
and so
\begin{equation}
    \begin{aligned}
        D(A_0+\epsilon Z)^2
        &\approx \epsilon^2\sum_{ijk\ell} H^{ij,k\ell} Z_{ij} Z_{k\ell} 
    \end{aligned}
\end{equation}

>>>>

Let two points on $\allS$ be $x_1$ and $x_2$, let $\bar x = (x_1+x_2)/2$, and let $z=(x_2 - x_1)/2$.
Then with $\grad D^2$ and $\grad^2 D^2$ the first and second derivatives of $D^2$, respectively,
Taylor expanding about $x_1$ and $x_2$ finds that
\begin{align*}
    D(\bar x)^2 
    &= D(x_1)^2 + \grad D(x_1)^2 \cdot z + \frac{1}{2} z^T \grad^2  D(x_1)^2 z + O(\|z\|^3) \\
    &=  D(x_2)^2 - \grad D(x_2)^2 \cdot z + \frac{1}{2} z^T \grad^2  D(x_2)^2 z + O(\|z\|^3) .
\end{align*}
Now, since $ D(x_1)^2 =  D(x_2)^2 = \grad D(x_1)^2 = \grad D(x_2)^2 = 0$ and
\begin{align*}
    \grad D(x_2)^2 = \grad D(x_1)^2 + 2 z^T \grad^2  D(x_1)^2 + O(\|z\|^2), \quad \text{and} \\
    \grad^2 D(x_2)^2 = \grad^2 D(x_1)^2 + O(\|z\|), 
\end{align*}
adding together the two equations above and dividing by two gets that
\begin{align*}
     D(\bar x)^2
    &= - \frac{3}{2} z^T \grad^2  D(x_1)^2 z + O(\|z\|^3) .
\end{align*}

@@

We could use this to get the quadratic approximation to $D$ near the optimal set.
To do so, it'd be nice to have a way to compute the inner integral above.
Suppose that we diagonalize $A = U \Lambda U^{-1}$.
Then
\begin{equation} \label{eqn:exp_deriv}
  \begin{aligned}
      \int_0^t e^{As} Z e^{A(t-s)} ds 
      &=
      \int_0^t U e^{\Lambda s} U^{-1} Z U e^{\Lambda (t-s)} U^{-1} ds 
  \end{aligned}
\end{equation}
Now, notice that
\begin{equation}
  \begin{aligned}
      \int_0^t e^{s \lambda_i} e^{(t-s) \lambda_j} ds
      &=
      \frac{ e^{t \lambda_i} - e^{t \lambda_j} }{ \lambda_i - \lambda_j } 
          \qquad & \text{if} \quad i \neq j \\
      &=
          t e^{t \lambda_i} 
          \qquad & \text{if} \quad i = j 
  \end{aligned}
\end{equation}
Therefore, 
defining
\begin{equation}
    \begin{aligned}
    X_{ij}(t,Z) 
       &= 
        \left( U^{-1} Z U \right)_{ij}
      \frac{ e^{t \lambda_i} - e^{t \lambda_j} }{ \lambda_i - \lambda_j } 
          \qquad & \text{if} \quad i \neq j \\
      &=
          \left( U^{-1} Z U \right)_{ii}
          t e^{t \lambda_i} 
          \qquad & \text{if} \quad i = j 
    \end{aligned}
\end{equation}
moving the $U$ and $U^{-1}$ outside the integral and integrating we get that
\begin{equation}
  \begin{aligned}
      \int_0^t e^{As} Z e^{A(t-s)} ds 
      &=
      U X(t,Z) U^{-1} .
  \end{aligned}
\end{equation}
This implies that
\begin{equation}
    \begin{aligned}
        D(A_0+\epsilon Z)^2
        &\approx \frac{1}{2} \epsilon^2 
        \int_0^\infty
            \rho(t) \tr\left\{ C U X(t,Z) U^{-1} B B^T (U^{-1})^T X(t,Z)^T U^T C^T \right\}
        dt .
    \end{aligned}
\end{equation}

To compute the $n^2 \times n^2$ matrix $H$,
we see that if $Z=\Delta_{k \ell}$, then
\begin{equation}
  \begin{aligned}
      X_{ij}^{k\ell}(t) 
      &= 
      (U^{-1})_{\cdot k} U_{\ell \cdot}
      \frac{ e^{t \lambda_i} - e^{t \lambda_j} }{ \lambda_i - \lambda_j } 
          \qquad & \text{if} \quad i \neq j \\
      &=
      (U^{-1})_{\cdot k} U_{\ell \cdot}
      t e^{t \lambda_i} 
          \qquad & \text{if} \quad i = j 
  \end{aligned}
\end{equation}
where $U_{k \cdot}$ is the $k$th row of $U$,
and so
\begin{equation}
    \begin{aligned}
        H_{ij, k\ell}(A)
        &=
        \int_0^\infty
            \rho(t) \tr\left\{ C U X^{ij}(t) U^{-1} B B^T (U^{-1})^T X^{k\ell}(t)^T U^T C^T \right\}
        dt .
    \end{aligned}
\end{equation}
This implies that
\begin{equation}
    \begin{aligned}
        D(A_0+\epsilon Z)^2
        &\approx \frac{1}{2} \epsilon^2 \sum_{ijk\ell} H_{ij,k\ell}(A_0) Z_{ij} Z_{k\ell}  .
    \end{aligned}
\end{equation}
